Learning for Logistic Regression Model in Parallel

### 2013-09-06 draft-backup.tex
revised once

% This is paper draft
% edited by Haoruo Peng
% Date: 2013.6.21
\documentclass[10pt, conference, compsocconf]{IEEEtran}
\usepackage{amsmath, amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{bbding}
\usepackage{epstopdf}

@@

\author{
\IEEEauthorblockN{Haoruo Peng}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\penghaoruo@hotmail.com}
\and
\IEEEauthorblockN{Ding Liang}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Ding\_Liang@htc.com}
\and
\IEEEauthorblockN{Deli Zhao}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\zhaodeli@gmail.com}
\and
\IEEEauthorblockN{Cyrus Choi}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Cyrus\_Choi@htc.com}
\and
\IEEEauthorblockN{Edward Y. Chang}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Edward\_Chang@htc.com}
}

@@

\begin{abstract}
Logistic regression (LR) model has various applications in machine learning.
LR is a supervised learning model for pattern classification and enjoys the benefits of statistical principles.
This paper addresses the issue of computational efficiency for solving LR in the big data scenario.
We compare the performance between different computing platforms and different parallel algorithms.
Our goal is to provide valuable suggestions on choosing computing platforms and learning algorithms according to different application situations for specific datasets and machine resources.
Three different parallel computing platforms along with three different types of parallel algorithms are carefully studied in this paper.
In addition, we present a novel parallel sublinear method to optimize the solution of LR.
%Hadoop, Mahout and Spark are three existing parallel computing platforms that can be implemented with large scale machine learning algorithms.
%Parallel gradient Descent and stochastic gradient descent are two state-of-art algorithms for solving LR.
To enhance capability, we consider the tradeoff between efficiency and accuracy. Extensive experimental results verify the effective performance of proposed algorithms.
\end{abstract}

@@

In recent years, many modern datasets grow drastically in both data volume and data dimensionality.
Large data volume and high data dimensionality bring unavoidable computational challenges to machine learning problems.
For example, in social networks, we have datasets~\cite{lewis2008tastes} covering multiple dimensions like friendship relationship and share of common interests. They generally include millions of records over millions of attributes. More evidently, for the task of text and multimedia categorization, we usually have to handle datasets with data volumes in billion scales and each data instance is characterized by thousands of feature dimensions~\cite{chang2011foundations}.
We tackle these challenges on both algorithm level and computing platform level via parallelization methods.
For many applications, in order to attain speedup in performance, we can employ not only task parallelism, but also data parallelism~\cite{subhlok1993exploiting}.
In this paper, we mainly discuss issues in task parallelism.

When developing parallel algorithms for LR, it is inevitable to come across the question of choosing which parallel computing platform to employ.
After a thorough survey, we choose three unique and best-known computing platforms to test: Hadoop~\cite{white2012hadoop}, Mahout~\cite{mahoutscalable} and Spark~\cite{zaharia2010spark}.
Hadoop employs HDFS~\cite{borthakur2008hdfs} and MapReduce~\cite{dean2008mapreduce}.
Mahout runs on Apache Hadoop and makes improvements in low level code.
Spark promotes efficiency of iterative algorithms and also supports HDFS.
We will compare design features of these three computing platforms separately in Section~\ref{sec:platform}.

@@

\subsection{Scalability Solutions}
%
In big data scenarios, large datasets require us to develop machine learning algorithms towards a more efficient and more parallelized end.
Researchers have already done much work to cater for the needs of massive datasets.
Early work like Parallel Support Vector Machines (PSVM)~\cite{chang2011psvm} employed an approximation matrix decomposition method based on row selections to reduce memory usage.
It can then increase the number of parallel computational nodes to several hundreds.
Later, the work of Parallel Latent Dirichlet Allocation (PLDA)~\cite{Liu:2011:PPL:1961189.1961198}~\cite{newman2007distributed} improves the computational efficiency of the LDA model by means of sampling methods.
The parallel algorithm in Hadoop is robust, for it has fault tolerance of machine failures, thus taking advantage of Hadoop features.
Recently, Dean \etal~\cite{le2011building} claimed that the advantage of parallelization is fully taken in deep learning algorithms.
It pushed the limit of parallel computational nodes to a hundred-million level, and in the meantime, achieved the best learning performance ever.
GraphLab~\cite{kyrola2012graphchi} is a recently developed tool for large-scale machine learning. It tackles efficiency and scalability problems of machine leaning algorithms by graph-theoretical models.

\subsection{Computing Platforms}\label{sec:platform}
In this paper, we specifically compares parallelization efforts of LR in two dimensions: platform and algorithm.

Three computing platforms that we are working on have their own advantages to implement machine learning algorithms.
The Apache Hadoop~\cite{white2012hadoop} software library allows for distributed processing of large datasets across clusters of computers using simple programming models.
Hadoop utilizes MapReduce~\cite{dean2008mapreduce} as its computational paradigm.
MapReduce paradigm takes a set of key-value pairs as its input.
Then a user-defined \textit{map} function turns the input into some intermediate key-value pairs.
Another user-defined \textit{reduce} function merges values for the same intermediate pairs to produce output.
All of the operations are independent, thereby making MapReduce paradigm easy to be parallelized.
In addition, Hadoop provides a Distributed File System (HDFS).
Both MapReduce and HDFS are designed to handle node failures in an automatic way, making Hadoop support large clusters that are built on commodity hardware.
The goal of Apache Mahout~\cite{mahoutscalable} is to build scalable machine learning libraries. Their core algorithms for clustering, classification and collaborative filtering are implemented on top of Apache Hadoop using the MapReduce paradigm. The core libraries are highly optimized so that Apache Mahout also has good performance for non-distributed algorithms.
Spark~\cite{zaharia2010spark} is a cluster computing system that aims to make data analysis fast.
Spark supports in-memory cluster computing. A job can load data into memory and query it repeatedly by creating and caching resilient distributed datasets (RDDs).
Moreover, RDDs achieve fault tolerance through lineage: information about how RDDs are derived from other RDDs is stored in reliable storage, thus making RDDs easy to be rebuilt if a certain partition is lost.
Spark can run up to 100 times faster than Hadoop for iterative algorithms.

In Table~\ref{tab:systemcmp}, we compare different features between Hadoop and Spark computing platforms.
As Mahout is mainly built on Hadoop, it inherits most of Hadoop features.
For simplicity, we use short abbreviations in the table.
Spark supports two types of operations on RDDs: Actions and Transformations (As and Ts).
Actions include functions like \textit{count}, \textit{collect} and \textit{save}. They usually return a result from input RDDs.
Transformations include functions like \textit{map}, \textit{filter} and \textit{join}. They normally build new RDDs from other RDDs, which comply with the lineage rule.
In the table, we use ``NFS" to represent ``Normal File System" and "Fault Tolerance" here is specifically for node failure situations.
\begin{table}[h]
\centering
\caption{Platform Comparison: Hadoop vs. Spark}\label{tab:systemcmp}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|}
\hline
           & Hadoop & Spark \\
\hline
Computational Paradigm  & MapReduce & As and Ts \\
\hline
File System Supported & HDFS & HDFS and NFS \\
\hline
Design Concept  & Key-value Pairs & RDD \\
\hline
Fault Tolerance Technique & Redundancy & Lineage \\
\hline
\end{tabular}\vspace{-0.5cm}
\end{table}

\subsection{Sublinear Methods}
Sublinear methods are recently proposed by many researchers.
Clarkson \etal~\cite{clarkson2010sublinear} first presented the solution of approximation algorithms in sublinear time.
It is applied to optimization problems for linear classifiers and minimum enclosing balls.
The algorithm employs a novel sampling technique along with a new multiplicative update procedure.
Hazan \etal~\cite{hazanbeating} exploited the approach to linear SVM.
Cotter \etal~\cite{cotter2012kernelized} extended the method to kernelized SVM.
%In~\cite{hazan2011optimal}, Hazan \etal applied the sublinear approximation approach to solve linear regression with penalties.
Garber and Hazan~\cite{garberapproximating} also developed it in Semi-Definite Programming (SDP).
Peng \etal~\cite{peng2012sublinear} utilized the method in the LR model with penalties and developed sequential sublinear algorithms for both $\ell_1$-penalty and $\ell_2$-penalty.

@@

The learning process aims to optimize $\w$ and $b$.
To make the optimization problem more practical to solve, a common practice is to add penalties to LR model.
Typically, for $\ell_2$-penalty, we solve the following optimization problem:\vspace{-0.2cm}
\begin{equation} \label{eqn:1}
	\max_{\bw ,b} \; \Big\{F(\w, b| \XM) - \frac{\lambda}{2}  \|\bw\|_2^{2} \Big\}.\vspace{-0.2cm}
\end{equation}
For $\ell_1$-penalty, we optimize the following criterion: \vspace{-0.2cm}
\begin{equation} \label{eqn:2}
	\max_{\bw ,b} \;  \big\{F(\w, b|\XM)- \gamma{\|\bw \|}_{1}\big\}.\vspace{-0.2cm}	
\end{equation}
Here, we define $F(\w, b| \XM) = \sum_{i=1}^n \log g_i(y_i)$ in both (\ref{eqn:1}) and (\ref{eqn:2}). To be brief, we omit the derivation here.

@@

In Algorithm~1, we give the sequential sublinear approximation procedure for logistic regression. The algorithm is mainly from~\cite{peng2012sublinear}. Symbols we use here comply with the definition in Section~\ref{sec:def}. $\varepsilon$ controls learning rate, $T$ determines iteration number, and $\eta$ defines approximation level.
	\begin{table} [ht]
	\begin{tabular}{l}
	\hline\noalign{\smallskip}
	\textbf{Algorithm 1}: Sub-Linear Logistic Regression (SLLR) \\
	\noalign{\smallskip}
	\hline
	\noalign{\smallskip}
	1:	Input parameters: $\varepsilon, \nu~or~\gamma, X, Y$ \\
	2:	Initialize parameters: $T, \eta, {\mathbf{u}}_{0}, {\bw}_{1}, {\mathbf{\bq}}_{1}, {b}_{1}$\\
	3:	Iterations: $t=1 \sim T$ \\
	4:	\tspace ${\bp}_{t}\leftarrow{\bq}_{t}/{\|{\bq}_{t}\|}_{1}$ \\
	5:  \tspace Choose ${i}_{t}\leftarrow i$ with probability $\bp(i)$ \\
	6:	\tspace $coef\leftarrow{y}_{{i}_{t}}g\lc-{y}_{{i}_{t}}\lc {{\bw}_{t}}^{T}{\mathbf{x}}_{i_t}+{b}_{t} \rc\rc$ \\
	7:	\tspace ${\bu}_{t}\leftarrow {\bu}_{t-1}+\frac{coef}{\sqrt{2T}}{\mathbf{x}}_{{i}_{t}}$ \\
    8:  \tspace ${\bxi}_{t}\leftarrow \argmax_{\bxi\in \Lambda}\lc{{\bp}_{t}}^{T}\bxi\rc$, if input $\nu$ for $\ell_2$-penalty \\
    9:  \tspace $\bu_t\leftarrow$ soft-thresholding operations, if input $\gamma$ for $\ell_1$-penalty \\
	10:	\tspaces ${\bw}_{t}\leftarrow {\bu}_{t}/\max \left\{1,\|{\bu}_{t}\|_2 \right\}$ \\
    11: \tspaces ${b}_{t}\leftarrow \sgn\lc {{\bp}_{t}}^{T}\mathbf{y}\rc$ \\
	12:	\tspaces Choose ${j}_{t}\leftarrow j$ with probability ${{\bw}_{t}\lj}^{2}/{\|{\bw}_{t}\|_2}^{2} $ \\
	13:	\tspaces Iterations: $i=1 \sim n$ \\
	14:	\tspaces\tspace $\sigma \leftarrow \mathbf{x}_{i} \lc {j}_{t}\rc{\|{\bw}_{t}\|_2}^{2}/{\bw}_{t}\lc {j}_{t} \rc+{\bxi}_{t}\li+{y}_{i}{b}_{t}$ \\
	15:	\tspaces\tspace $\hat{\sigma} \leftarrow clip\lc \sigma,1/\eta \rc$ \\
	16:	\tspaces\tspace ${\bq}_{t+1}\li \leftarrow {\bp}_{t}\li\lc 1-\eta\hat{\sigma} + {\eta}^{2}{\hat{\sigma}}^{2} \rc$ \\
	17:	Output: $\bar{\bw}=\frac{1}{T}\sum_{t}{\bw}_{t},\bar{b}=\frac{1}{T}\sum_{t}{b}_{t}$ \\
	\hline
	\end{tabular} 	
	\label{alg:1}
	\end{table}
Each iteration of the SLLR algorithm has two phases: \textit{stochastic primal update} and \textit{stochastic dual update}.
Steps from line 4 to line 11 consist of the primal part while the dual part is composed of steps from line 12 to line 16.
We give the sublinear algorithm in a unified way for $\ell_2$-penalty and $\ell_1$-penalty.
If we are dealing with $\ell_2$-penalty, we ignore line 9 and accomplish the computation in line 8 by a simple greedy algorithm.
Here, $\Lambda$ represents a Euclidean space with conditions $
\Lambda=\left\{ \bxi\in\mathbb{R}_{n} \, | \, \forall i, \, 0\leq{\xi}_{i}\leq 2, \, {\|\xi \|}_{1}\leq \nu n  \right
\}$.
If we are faced with $\ell_1$-penalty, we ignore line 8 and expand the procedure of line 9 as following.
\begin{table} [ht]
\begin{tabular}{l}
\hline\noalign{\smallskip}
\textbf{Procedure}: Line 9 in Algorithm 1\\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
Iterations: $j=1 \sim d$ \\
\tspace \textbf{if} $\buprev_t\lj>0$ \textbf{and} $\bu_t\lj>0$ \\
\tspace\tspace $\bu_t\lj=\max \lc \bu_t\lj-\gamma ,0 \rc$ \\
\tspace \textbf{if} $\buprev_t\lj<0$ \textbf{and} $\bu_t\lj<0$ \\
\tspace\tspace $\bu_t\lj=\min \lc \bu_t\lj+\gamma ,0 \rc$ \\
$\buprev_{t+1}\leftarrow\bu_t$ \\
\hline
\end{tabular}
\end{table}

@@

    21: \tspaces $\bp_{t+1}$$\leftarrow$Dual-Update($\bp_t$) \\

@@

There are also three more things that need to be handled here.
First is the parameter passing issue. It is critical to choose an efficient way to pass the updated parameters between iterations and even between different MapReduce jobs.
It is even more challenging when we have to deal with HDFS. It is clear that the fastest communication way is to pass parameters by \textit{Configuration()}.
However, it always suffers from memory buffer limits, and will greatly impair scalability when the number of parameters is huge.
Another way is to compress the parameter sequence in a string when sending and uncompress them when receiving. It has computational overhead and it still cannot avoid the issue of memory buffer limit.
Considering the philosophy of the design of Hadoop, the natural thing to do is to pass parameters by file, the same way to pass all the data.
Although such an operation is insufficiently efficient, it is the feasible way to support large datasets.

@@

The datasets are generally sparse when data dimensionally is high. This characteristic makes us focus on dealing with data sparsity issue when writing code.
This is the third important thing for the algorithm, although it is not reflected in the pseudo-code.
Instead of naively writing the simple code in data intensive situations, we assign the \textit{index} for each data value in the sparse vector, and all computations are changed accordingly.
This sparse format brings us great efficiency improvement, and all results in Section~\ref{sec:experiment} are obtained from the code for sparsity.

\subsection{Parallel Sublinear algorithms in Spark}
The algorithm to solve sublinear learning for penalized logistic regression in Spark is shown below.
In the pseudo-code of Algorithm 3, the procedure for Primal-Update and the procedure for Dual-Map are the same as those in Algorithm PSUBPLR-MR.
    \begin{table}[ht]
	\begin{tabular}{l}
    \hline\noalign{\smallskip}
	\textbf{Algorithm 3}: PSUBPLR-SPARK \\
	\noalign{\smallskip}
	\hline
	\noalign{\smallskip}
    1:  ~Input parameters: $\varepsilon, \nu~or~\gamma, X, Y, n, d$ \\
    2:	~Initialize parameters: $T, \eta, {\mathbf{u}}_{0}, {\bw}_{1}, {\mathbf{\bq}}_{1}, {b}_{1}$\\
    3:  ~points $\leftarrow$ spark.textFile(inputfile).map(parsePoint()).cache() \\
    4:  ~Iterations: $t=1 \sim T$ \\
    5:  ~\tspace gradient $\leftarrow$ points.map($(\frac{1}{1+e^{-y((\bw_t^T \textbf{x})+b)}}-1) * y * \bp [index]$) \\
        ~~~\tspace\tspace\tspace\tspace\tspace .reduce(sum()) \\
    6:  ~\tspace ($\bw_{t+1}, b_{t+1}$)$\leftarrow$Primal-Update($\bw_t, b_t$) \\
    7:  ~\tspace Choose $j_t \leftarrow j$ with probability ${{\bw}_{t+1}\lj}^{2}/{\|{\bw}_{t+1}\|_2}^{2} $ \\
    8:  ~\tspace pAdjust $\leftarrow$ points.map(MW-Update()).reduce(copy()) \\
    9:  ~\tspace $\bp_{t+1}$$\leftarrow$Dual-Update($\bp_t$)  \\
    10: \textbf{Output}($\bw, b$) \\
    \hline
    \end{tabular}
	\end{table}

@@

In parallel mode, the primal update contains the update of $\bw_t$, which takes $O(n)$ time.
The dual update contains the $\ell_2$-sampling process for the choice of $j_t$ in $O(d)$ time, and the update of $\bp$ in $O(1)$ time.
Altogether, each iteration takes $O(n+d)$ time.
Compared to the analysis of sequential algorithm, parallelization does not necessarily change computational complexity.
Theoretically, by deeper analysis of constant coefficients of the obtained computational complexity, parallel sublinear algorithm can be two times faster than the sequential version.
Moreover, by starting two separate MapReduce jobs in one iteration simultaneously, the running time can be reduced to $O(max\{n,d\})$.

>>>>

% This is paper draft
% edited by Haoruo Peng
% Date: 2013.9.5
\documentclass[10pt, conference, compsocconf]{IEEEtran}
\usepackage{amsmath, amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{bbding}
\usepackage{epstopdf}

@@

\author{
\IEEEauthorblockN{Haoruo Peng}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\penghaoruo@hotmail.com}
\and
\IEEEauthorblockN{Ding Liang}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Ding\_Liang@htc.com}
\and
%\IEEEauthorblockN{Deli Zhao}
%\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\zhaodeli@gmail.com}
%\and
\IEEEauthorblockN{Cyrus Choi}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Cyrus\_Choi@htc.com}
\and
\IEEEauthorblockN{Edward Y. Chang}
\IEEEauthorblockA{HTC Research Lab\\Beijing, China\\Edward\_Chang@htc.com}
}

@@

\begin{abstract}
Logistic regression (LR) model has various applications in machine learning.
LR is a supervised learning model for pattern classification and enjoys the benefits of statistical principles.
This paper addresses the issue of computational efficiency for solving LR in big data scenario.
We compare the performance between different computing platforms and different parallel algorithms.
Our goal is to provide valuable suggestions on choosing computing platforms and learning algorithms according to different application situations for specific datasets and machine resources.
We also present a novel parallel sublinear method to optimize the solution of LR.
Overall, two different parallel computing platforms along with three different types of parallel algorithms are carefully studied in this paper.
To enhance capability, we consider the tradeoff between efficiency and accuracy.
Extensive experimental results verify the effective performance of proposed algorithms.
\end{abstract}

@@

In recent years, many modern datasets grow drastically in both data volume and data dimensionality.
Large data volume and high data dimensionality bring unavoidable computational challenges to machine learning problems.
For example, in social networks, we have datasets covering multiple dimensions like friendship relationship and share of common interests. They generally include millions of records over millions of attributes. More evidently, for the task of text and multimedia categorization, we usually have to handle datasets with data volumes in billion scales and each data instance is characterized by thousands of feature dimensions.
We tackle these challenges on both algorithm level and computing platform level via parallelization methods.
For many applications, in order to attain speedup in performance, we can employ not only task parallelism, but also data parallelism~\cite{subhlok1993exploiting}.
In this paper, we mainly discuss issues in task parallelism.

When developing parallel algorithms for LR, it is inevitable to come across the question of choosing which parallel computing platform to employ.
After a thorough survey, we choose two unique and best-known computing platforms to test: Hadoop~\cite{white2012hadoop} and Spark~\cite{zaharia2010spark}.
Hadoop employs HDFS~\cite{borthakur2008hdfs} and MapReduce~\cite{dean2008mapreduce}.
Spark promotes efficiency of iterative algorithms and also supports HDFS.
We will compare design features of these two computing platforms in Section~\ref{sec:platform}.

@@

%\subsection{Scalability Solutions}
%
%In big data scenarios, large datasets require us to develop machine learning algorithms towards a more efficient and more parallelized end. Researchers have already done much work to cater for the needs of massive datasets. Early work like Parallel Support Vector Machines (PSVM)~\cite{chang2011psvm} employed an approximation matrix decomposition method based on row selections to reduce memory usage. It can then increase the number of parallel computational nodes to several hundreds. Later, the work of Parallel Latent Dirichlet Allocation (PLDA)~\cite{Liu:2011:PPL:1961189.1961198}~\cite{newman2007distributed} improves the computational efficiency of the LDA model by means of sampling methods. The parallel algorithm in Hadoop is robust, for it has fault tolerance of machine failures, thus taking advantage of Hadoop features. Recently, Dean \etal~\cite{le2011building} claimed that the advantage of parallelization is fully taken in deep learning algorithms. It pushed the limit of parallel computational nodes to a hundred-million level, and in the meantime, achieved the best learning performance ever. GraphLab~\cite{kyrola2012graphchi} is a recently developed tool for large-scale machine learning. It tackles efficiency and scalability problems of machine leaning algorithms by graph-theoretical models.

In this paper, we specifically compares parallelization efforts of LR in two dimensions: platform and algorithm.
\subsection{Computing Platforms}\label{sec:platform}
Two computing platforms that we are working on have their own advantages to implement machine learning algorithms.
The Hadoop~\cite{white2012hadoop} platform allows for distributed processing of large datasets across clusters of computers using simple programming models. It utilizes MapReduce~\cite{dean2008mapreduce} as its computational paradigm, which is easy to be parallelized. In addition, Hadoop provides a Distributed File System (HDFS). Both MapReduce and HDFS are designed to handle node failures in an automatic way, making Hadoop support large clusters that are built on commodity hardware.
Spark~\cite{zaharia2010spark} is a cluster computing system that aims to make data analysis fast. It supports in-memory cluster computing. A job can load data into memory and query it repeatedly by creating and caching resilient distributed datasets (RDDs). Moreover, RDDs achieve fault tolerance through lineage: information about how RDDs are derived from other RDDs is stored in reliable storage, thus making RDDs easy to be rebuilt if a certain partition is lost. Spark can run up to 100 times faster than Hadoop for iterative algorithms.

In Table~\ref{tab:systemcmp}, we compare different features between Hadoop and Spark computing platforms.
For simplicity, we use short abbreviations in the table.
Spark supports two types of operations on RDDs: Actions and Transformations (As and Ts).
Actions include functions like \textit{count}, \textit{collect} and \textit{save}. They usually return a result from input RDDs.
Transformations include functions like \textit{map}, \textit{filter} and \textit{join}. They normally build new RDDs from other RDDs, which comply with the lineage rule.
In the table, we use ``NFS" to represent ``Normal File System" and "Fault Tolerance" here is specifically for node failure situations.
\begin{table}[h]
\centering
\caption{Platform Comparison: Hadoop vs. Spark}\label{tab:systemcmp}\vspace{-0.3cm}
\begin{tabular}{|c|c|c|}
\hline
           & Hadoop & Spark \\
\hline
Computational Paradigm  & MapReduce & As and Ts \\
\hline
File System Supported & HDFS & HDFS and NFS \\
\hline
Design Concept  & Key-value Pairs & RDD \\
\hline
Fault Tolerance Technique & Redundancy & Lineage \\
\hline
\end{tabular}\vspace{-0.5cm}
\end{table}

It is worth mentioning that Apache Mahout~\cite{mahoutscalable} runs on Hadoop, and it is a scalable machine learning library. It integrates many core algorithms for clustering, classification and collaborative filtering in its package. It is highly optimized so that Mahout also has good performance for non-distributed algorithms. We will use it as a baseline in the experiments.

\subsection{Sublinear Methods}
Sublinear methods are recently proposed by many researchers.
Clarkson \etal~\cite{clarkson2010sublinear} first presented the solution of approximation algorithms in sublinear time.
It is applied to optimization problems for linear classifiers and minimum enclosing balls.
The algorithm employs a novel sampling technique along with a new multiplicative update procedure.
Hazan \etal~\cite{hazanbeating} exploited the approach to linear SVM.
Cotter \etal~\cite{cotter2012kernelized} extended the method to kernelized SVM.
%In~\cite{hazan2011optimal}, Hazan \etal applied the sublinear approximation approach to solve linear regression with penalties.
%Garber and Hazan~\cite{garberapproximating} also developed it in Semi-Definite Programming (SDP).
Peng \etal~\cite{peng2012sublinear} utilized the method in the LR model with penalties and developed sequential sublinear algorithms for both $\ell_1$-penalty and $\ell_2$-penalty.

@@

The learning process aims to optimize $\w$ and $b$.
It is a common practice to add penalties to LR model in order to avoid overfitting and make the optimization result more practical.
Typically, for $\ell_2$-penalty, we solve the following optimization problem:\vspace{-0.2cm}
\begin{equation} \label{eqn:1}
	\max_{\bw ,b} \; \Big\{F(\w, b| \XM) - \frac{\lambda}{2}  \|\bw\|_2^{2} \Big\}.\vspace{-0.2cm}
\end{equation}
For $\ell_1$-penalty, we optimize the following criterion: \vspace{-0.2cm}
\begin{equation} \label{eqn:2}
	\max_{\bw ,b} \;  \big\{F(\w, b|\XM)- \gamma{\|\bw \|}_{1}\big\}.\vspace{-0.2cm}	
\end{equation}
Here, we define $F(\w, b| \XM) = \sum_{i=1}^n \log g_i(y_i)$ in both (\ref{eqn:1}) and (\ref{eqn:2}). To be brief, we omit the derivation here.

@@

In Algorithm~1, we give the sequential sublinear approximation procedure for logistic regression. The algorithm is mainly from~\cite{peng2012sublinear}. Symbols we use here comply with the definition in Section~\ref{sec:def}. $\varepsilon$ controls learning rate, $T$ determines iteration number, and $\eta$ defines approximation level. Details of the parameter initialization step is studied in~\cite{peng2012sublinear}. To be concise, we omit them in this paper.
	\begin{table} [ht]
	\begin{tabular}{l}
	\hline\noalign{\smallskip}
	\textbf{Algorithm 1}: Sub-Linear Logistic Regression (SLLR) \\
	\noalign{\smallskip}
	\hline
	\noalign{\smallskip}
	1:	Input parameters: $\varepsilon, \nu~or~\gamma, X, Y$ \\
	2:	Initialize parameters: $T, \eta, {\mathbf{u}}_{0}, {\bw}_{1}, {\mathbf{\bq}}_{1}, {b}_{1}$\\
	3:	Iterations: $t=1 \sim T$ \\
	4:	\tspace ${\bp}_{t}\leftarrow{\bq}_{t}/{\|{\bq}_{t}\|}_{1}$ \\
	5:  \tspace Choose ${i}_{t}\leftarrow i$ with probability $\bp(i)$ \\
	6:	\tspace $coef\leftarrow{y}_{{i}_{t}}g\lc-{y}_{{i}_{t}}\lc {{\bw}_{t}}^{T}{\mathbf{x}}_{i_t}+{b}_{t} \rc\rc$ \\
	7:	\tspace ${\bu}_{t}\leftarrow {\bu}_{t-1}+\frac{coef}{\sqrt{2T}}{\mathbf{x}}_{{i}_{t}}$ \\
    8:  \tspace ${\bxi}_{t}\leftarrow \argmax_{\bxi\in \Lambda}\lc{{\bp}_{t}}^{T}\bxi\rc$, if input $\nu$ for $\ell_2$-penalty \\
    9:  \tspace $\bu_t\leftarrow$ soft-thresholding operations, if input $\gamma$ for $\ell_1$-penalty \\
	10:	\tspaces ${\bw}_{t}\leftarrow {\bu}_{t}/\max \left\{1,\|{\bu}_{t}\|_2 \right\}$ \\
    11: \tspaces ${b}_{t}\leftarrow \sgn\lc {{\bp}_{t}}^{T}\mathbf{y}\rc$ \\
	12:	\tspaces Choose ${j}_{t}\leftarrow j$ with probability ${{\bw}_{t}\lj}^{2}/{\|{\bw}_{t}\|_2}^{2} $ \\
	13:	\tspaces Iterations: $i=1 \sim n$ \\
	14:	\tspaces\tspace $\sigma \leftarrow \mathbf{x}_{i} \lc {j}_{t}\rc{\|{\bw}_{t}\|_2}^{2}/{\bw}_{t}\lc {j}_{t} \rc+{\bxi}_{t}\li+{y}_{i}{b}_{t}$ \\
	15:	\tspaces\tspace $\hat{\sigma} \leftarrow clip\lc \sigma,1/\eta \rc$ \\
	16:	\tspaces\tspace ${\bq}_{t+1}\li \leftarrow {\bp}_{t}\li\lc 1-\eta\hat{\sigma} + {\eta}^{2}{\hat{\sigma}}^{2} \rc$ \\
	17:	Output: $\bar{\bw}=\frac{1}{T}\sum_{t}{\bw}_{t},\bar{b}=\frac{1}{T}\sum_{t}{b}_{t}$ \\
	\hline
	\end{tabular} 	
	\label{alg:1}
	\end{table}
Each iteration of the SLLR algorithm has two phases: \textit{stochastic primal update} and \textit{stochastic dual update}.
Steps from line 4 to line 11 consist of the primal part while the dual part is composed of steps from line 12 to line 16.
We give the sublinear algorithm in a unified way for $\ell_2$-penalty and $\ell_1$-penalty.
If we are dealing with $\ell_2$-penalty, we ignore line 9 and accomplish the computation in line 8 by a simple greedy algorithm.
Here, $\Lambda$ represents a Euclidean space with conditions $
\Lambda=\left\{ \bxi\in\mathbb{R}_{n} \, | \, \forall i, \, 0\leq{\xi}_{i}\leq 2, \, {\|\xi \|}_{1}\leq \nu n  \right
\}$.
If we are faced with $\ell_1$-penalty, we ignore line 8 and expand the procedure of line 9 as following.
\begin{table} [ht]
\begin{tabular}{l}
\hline\noalign{\smallskip}
\textbf{Procedure}: Line 9 in Algorithm 1\\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
Iterations: $j=1 \sim d$ \\
\tspace \textbf{if} $\buprev_t\lj>0$ \textbf{and} $\bu_t\lj>0$ \\
\tspace\tspace $\bu_t\lj=\max \lc \bu_t\lj-\gamma ,0 \rc$ \\
\tspace \textbf{if} $\buprev_t\lj<0$ \textbf{and} $\bu_t\lj<0$ \\
\tspace\tspace $\bu_t\lj=\min \lc \bu_t\lj+\gamma ,0 \rc$ \\
$\buprev_{t+1}\leftarrow\bu_t$ \\
\hline
\end{tabular}
\end{table}

@@

    21: \tspaces $\bp_{t+1}\leftarrow$Dual-Update($\bp_t$) \\

@@

There are also three more things that need to be handled here.
First is the parameter passing issue. It is critical to choose an efficient way to pass the updated parameters between iterations and even between different MapReduce jobs. Considering the philosophy of the design of Hadoop, the natural thing to do is to pass parameters by file, the same way to pass all the data. We use a single file to pass all parameters each time when a job starts or finishes separately. Although such an operation is insufficiently efficient, it is the feasible way to support large datasets.

@@

The datasets are generally sparse when data dimensionally is high. This characteristic makes us focus on dealing with data sparsity issue when writing code. Instead of naively writing the simple code in data intensive situations, we only store pairs of \textit{index} and \textit{value} in a data vector. Then all computations are changed accordingly.
This sparse format brings us great efficiency improvement, and all results in Section~\ref{sec:experiment} are obtained from the code for sparsity.

\subsection{Parallel Sublinear algorithms in Spark}
The algorithm to solve sublinear learning for penalized logistic regression in Spark is shown below.
In the pseudo-code of Algorithm 3, the procedure for Primal-Update and the procedure for Dual-Map are the same as those in Algorithm PSUBPLR-MR.
    \begin{table}[ht]
	\begin{tabular}{l}
    \hline\noalign{\smallskip}
	\textbf{Algorithm 3}: PSUBPLR-SPARK \\
	\noalign{\smallskip}
	\hline
	\noalign{\smallskip}
    1:  ~Input parameters: $\varepsilon, \nu~or~\gamma, X, Y, n, d$ \\
    2:	~Initialize parameters: $T, \eta, {\mathbf{u}}_{0}, {\bw}_{1}, {\mathbf{\bq}}_{1}, {b}_{1}$\\
    3:  ~points $\leftarrow$ spark.textFile(inputfile).map(parsePoint()).cache() \\
    4:  ~Iterations: $t=1 \sim T$ \\
    5:  ~\tspace gradient $\leftarrow$ points.map($(\frac{1}{1+e^{-y((\bw_t^T \textbf{x})+b)}}-1) * y * \bp [index]$) \\
        ~~~\tspace\tspace\tspace\tspace\tspace .reduce(sum()) \\
    6:  ~\tspace ($\bw_{t+1}, b_{t+1}$)$\leftarrow$Primal-Update($\bw_t, b_t$) \\
    7:  ~\tspace Choose $j_t \leftarrow j$ with probability ${{\bw}_{t+1}\lj}^{2}/{\|{\bw}_{t+1}\|_2}^{2} $ \\
    8:  ~\tspace pAdjust $\leftarrow$ points.map(MW-Update()).reduce(copy()) \\
    9:  ~\tspace $\bp_{t+1}\leftarrow$Dual-Update($\bp_t$)  \\
    10: \textbf{Output}($\bw, b$) \\
    \hline
    \end{tabular}
	\end{table}

@@

In parallel mode, the primal update contains the update of $\bw_t$, which takes $O(n)$ time.
The dual update contains the $\ell_2$-sampling process for the choice of $j_t$ in $O(d)$ time, and the update of $\bp$ in $O(1)$ time.
Altogether, each iteration takes $O(n+d)$ time.
Compared to the analysis of sequential algorithm, parallelization does not necessarily change computational complexity.
Theoretically, parallel sublinear algorithm can be two times faster than the sequential version as the time for both update procedures in each iteraion is reduced to $O(1)$.
Moreover, by starting two separate MapReduce jobs in one iteration simultaneously, the running time can be reduced to $O(max\{n,d\})$.
